---
title: "Sample Data"
author: "Sabrina Heitmann"
date: "9/30/2020"
output: html_document
---

# In this code, I merged the metadata for the Hermiston and Corvallis genotypes to my uncorrected Sampledata (which has all of the info I sent to the CGRB). There were some discrepencies of genotype names between datasheets so I changed the Sample data Genotype names to reflect the metadata Genotype names. I then fixed any transcribing errors. 
* Note that BESC-unk (formally known as YK-unk) does not have any meta data associated with it

```{r include=FALSE}
library(tidyverse)
library(magrittr)
```

```{r}
meta <- read.csv("data/gardenData/GenotypeMetadata_new.csv",as.is=T)
sample_data <- read.csv("data/gardenData/SampleData_uncorrected.csv", as.is=T, fileEncoding = 'UTF-8-BOM') 
```

## Joined metadata and my sample data sheets together 
```{r}
new <- left_join(sample_data, meta, by = "Genotype")
```

## Identify the rows that are NA (meaning that there was no match in the metadata file) and group by genotype
Working backwards from NAs to change labels
```{r}
new_NA <- new[is.na(new$Latitude_New),]

new_NA %>% 
  group_by(Genotype) %>%
  count()
```

## Change "B-" to "BESC-" in sample datasheet
All of the BESC genotypes in sample_data are actually labeled "B-" so need to change to "BESC-" to match meta datasheet
Left join again
```{r}
sample_data <- sample_data %>%
  mutate(Genotype = gsub("B-", "BESC-", Genotype, fixed = TRUE))

new <- left_join(sample_data, meta, by = "Genotype")

new_NA <- new[is.na(new$Latitude_New),]

```

## Change "YK-", "CW-", "LS-" to "BESC-" and omit space from PCR_control in sample datasheet
Left join again
```{r}
sample_data <- sample_data %<>%
  mutate(Genotype = gsub("YK-", "BESC-", Genotype, fixed = TRUE)) %>%
  mutate(Genotype = gsub("CW-", "BESC-", Genotype, fixed = TRUE)) %>%
  mutate(Genotype = gsub("LS-", "BESC-", Genotype, fixed = TRUE)) %>% 
  mutate(Genotype = gsub("UNK", "unk", Genotype, fixed = TRUE)) %>%
  mutate(Genotype = gsub("PCR_control ", "PCR_control", Genotype, fixed = TRUE)) %>%
  mutate(Genotype = gsub("BESC-unk", "YK-unk", Genotype, fixed = TRUE))

new <- left_join(sample_data, meta, by = "Genotype")

new_NA <- new[is.na(new$Latitude_New),]
```

## Remove controls and BESC-unk only from new_NA datasheet to see which genotypes need to be fixed in SampleData

```{r}
new_NA <- new_NA %>% 
  filter(!grepl("Extract_control", Genotype))

new_NA <- new_NA %>% 
  filter(!grepl("Pos_control", Genotype))

new_NA <- new_NA %>% 
  filter(!grepl("PCR_control", Genotype))

new_NA <- new_NA %>% 
  filter(!grepl("YK-unk", Genotype))
    
```

## Sample_data Genotype changes
* "EWG_2019_Lib02_s128" - changed "-" to "BESC-833"
* "EWG_2019_Lib02_s171" - changed "-" to "YK-unk"
* "EWG_2019_Lib03_s270" - changed "unk" to "YK-unk"
* "EWG_2019_Lib01_s297" - changed "MCMN-27" to "MCMN-27-5"
* "EWG_2019_Lib01_s360" - changed "b-196" to "BESC-196"
* "EWG_2019_Lib02_s351" - changed "MCM-2" to "MCMN-27-5"
* "EWG_2019_Lib01_s478" - changed "Extraction_control" to "Extract_control"
* "EWG_2019_Lib01_s480" - changed "Positive_control" to "Pos_control"
* "EWG_2019_Lib02_s411" - changed "Positive_control" to "Pos_control"
```{r results = 'hide'}
new_NA[,1]

sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib02_s128"] <- "BESC-833")
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib02_s171"] <- "YK-unk")

#check to make sure both row's Genotype values have been changed - YES
which(sample_data$Genotype == "-")

sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib03_s270"] <- "YK-unk")
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib01_s297"] <- "MCMN-27-5")
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib01_s360"] <- "BESC-196")
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib02_s351"] <- "MCMN-27-5")
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib01_s478"] <- "Extract_control")
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib01_s480"] <- "Pos_control")
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib02_s411"] <- "Pos_control")

new <- left_join(sample_data, meta, by = "Genotype")

genotypes <- new %>%
  group_by(Genotype) %>%
  count()
```

## There are 60 genotypes so there still must be 1 wrong genotype name. I uploaded the Corvallis_longformat to compare the genotype names. Then I combined the coords to the new Sample
```{r results = 'hide'}

coords.corv <- read.csv("data/gardenData/Corvallis_longFormat.csv",as.is=T) %>% 
  mutate(Garden="Corvallis")

coords.herm <- read.csv("data/gardenData/Hermiston_longFormat.csv",as.is=T) %>% 
  mutate(Garden="Hermiston") %>% 
  filter(Region != c("Border"))

coords <- rbind(coords.corv, coords.herm)

coords %<>%
  mutate(Genotype = gsub("YK-", "BESC-", Genotype, fixed = TRUE)) %>%
  mutate(Genotype = gsub("CW-", "BESC-", Genotype, fixed = TRUE)) %>%
  mutate(Genotype = gsub("LS-", "BESC-", Genotype, fixed = TRUE)) %>% 
  mutate(Genotype = gsub("UNK", "unk", Genotype, fixed = TRUE)) %>%
  mutate(Genotype = gsub("PCR_control ", "PCR_control", Genotype, fixed = TRUE)) %>%
  mutate(Genotype = gsub("BESC-unk", "YK-unk", Genotype, fixed = TRUE))

# combine meta and coords datasheets to have the full sample information (includes samples that didn't have microbiome data collected)
all_tree_meta_data <- left_join(coords, meta, by = "Genotype")

all_tree_meta_data = mutate(all_tree_meta_data,
              ID = paste("C",Col, "R",Row,"B", Block, sep = ""))

new %>% filter(!Genotype %in% coords$Genotype)

sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib02_s407"] <- "BESC-273")

new <- left_join(sample_data, meta, by = "Genotype")

# Removed R17C29Hermiston and R3C37Corvallis because they had the same key (Row and Column) but were duplicated twice per garden
new = mutate(new,
              key = paste("C",C, "R",R,Garden,Genotype, sep = ""))

n_occur <- data.frame(table(new$key))
duplicate <- n_occur[n_occur$Freq > 1,]

sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib03_s015"),]
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib01_s301"),]
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib02_s129"),]
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib02_s171"),]

new <- left_join(sample_data, meta, by = "Genotype")

new = mutate(new,
              key = paste("C",C, "R",R, "B",Garden,Genotype, sep = ""))

n_occur <- data.frame(table(new$key))
duplicate <- n_occur[n_occur$Freq > 1,]
removed <- new[(new$key %in% duplicate$Var1),]
```

Then I combined the coords to the new merged datasheet and edited which variables I wanted to include in my metadata
* "EWG_2019_Lib02_s021" - Corvallis changed to Hermiston 
* "EWG_2019_Lib03_s040" - NA changed to 12
* "EWG_2019_Lib03_s072" - Corvallis changed to Hermiston
* "EWG_2019_Lib03_s209" - Hermiston changed to Corvallis
* "EWG_2019_Lib03_s227" - Hermiston changed to Corvallis
* "EWG_2019_Lib01_s342" - Corvallis changed to Hermiston
* "EWG_2019_Lib03_s353" - changed to Pos_control
```{r}
coords <- coords %>%  
    mutate(Row = as.numeric(Row)) %>% 
    mutate(Col = as.numeric(Col))

new <- new %>%  
    mutate(R = as.numeric(R)) %>% 
    mutate(C = as.numeric(C))

new = mutate(new,
              key = paste("C",C, "R",R, "B",Garden,Genotype, sep = ""))

coords = mutate(coords,
              key = paste("C",Col, "R",Row, "B",Garden,Genotype, sep = ""))

new <- left_join(new, coords, by = "key") 

# filters out NA's in Genotype.y and puts them in new data table
errors <- 
  new %>% 
  filter(is.na(Genotype.y)) %>%
  filter(!grepl("Extract_control", Genotype.x)) %>% 
  filter(!grepl("PCR_control", Genotype.x)) %>% 
  filter(!grepl("Pos_control", Genotype.x))

errors[,1]

#fixes errors
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib02_s021"] <- "Hermiston")
sample_data <- within(sample_data, R[Sample_ID=="EWG_2019_Lib03_s040"] <- "12")
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib03_s072"] <- "Hermiston")
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib03_s209"] <- "Corvallis")
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib03_s227"] <- "Corvallis")
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib01_s342"] <- "Hermiston")
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib03_s353"] <- "BESC-1207")
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib01_s080"] <- "Hermiston")
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib01_s087"] <- "Hermiston")
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib02_s002"] <- "Corvallis")
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib02_s049"] <- "Hermiston")
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib02_s053"] <- "Hermiston")
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib02_s147"] <- "BESC-1210") # from BESC-1201 (typo)
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib01_s342"] <- "Hermiston")
sample_data <- within(sample_data, Garden[Sample_ID=="EWG_2019_Lib01_s366"] <- "Corvallis")
sample_data <- within(sample_data, R[Sample_ID=="EWG_2019_Lib03_s355"] <- "14") #was R4 on plate datasheet (typo - missing the 1)
sample_data <- within(sample_data, C[Sample_ID=="EWG_2019_Lib01_s441"] <- "15") #was C5 on plate datasheet (typo - missing the 1)
sample_data <- within(sample_data, C[Sample_ID=="EWG_2019_Lib01_s442"] <- "7") #was C17 on sample_data (typo - added a 1)
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib01_s455"] <- "BESC-1201")
sample_data <- within(sample_data, Genotype[Sample_ID=="EWG_2019_Lib01_s456"] <- "BESC-286")

#There seems to be a sampling error from R2C37 to R4C37 in Corvallis so deleting all 3 from dataset
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib03_s181"),] #R4C37
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib03_s270"),] #R2C37
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib01_s291"),] #R4C37

#Remove duplicate R3C10 in Corvallis with wrong genotype
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib03_s301"),] #R4C37

# removed EWG_2019_Lib03_s353 and others because it's row col identity could not be corrected
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib03_s353"),]
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib03_s015"),]
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib01_s301"),]
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib02_s129"),]
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib02_s171"),]
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib01_s442"),]
sample_data<-sample_data[!(sample_data$Sample_ID=="EWG_2019_Lib01_s392"),]

new <- left_join(sample_data, meta, by = "Genotype") 

new <- new %>%  
    mutate(R = as.numeric(R)) %>% 
    mutate(C = as.numeric(C))

new = mutate(new,
              key = paste("C",C, "R",R, "B",Garden,Genotype, sep = ""))

new <- left_join(new, coords, by = "key") 

# Determine why there are still genotype missmatches - it is because of the controls
Genotype_error <- new %>% 
  select(Sample_ID, Genotype.x, Genotype.y) %>% 
  filter(!Genotype.x %in% Genotype.y)

# 

# remove unneeded columns
new <- new %>% 
  select(-c(Comment, Latitude, Longitude, Pop.1, Genotype.y, Garden.y, Number, Col, Row, key))

names(new)[7] <- "Genotype"
names(new)[8] <- "Garden"
names(new)[17] <- "Latitude"
names(new)[18] <- "Longitude"
```

Genotype, block, garden - should be the same with #
```{r}
write.csv(all_tree_meta_data, "data/gardenData/all_tree_meta_data.csv")
write.csv(new, "data/gardenData/Merged_meta&sampledata.csv", row.names = FALSE)
write.csv(coords, "data/gardenData/coords_genotypes_updated.csv", row.names = FALSE)
write.csv(sample_data, "data/gardenData/sample_data_corrected.csv", row.names = FALSE)
```





