---
title: "Correlation plots"
author: "Sabrina Heitmann"
date: "4/13/2022"
output: html_document
---
```{r}
library(vegan)
library(tidyverse)
library(phyloseq)
library(magrittr)
library(here)
library(ggthemes)
library(ggpubr)
```

Add iNEXT diversity and richness estimates to meta data and average estimates by genotype within garden
```{r}
phy.clean <- readRDS(here("EWG_github_2023", "output", "compiled", "phy.clean.rds"))

meta_data <- phy.clean %>% sample_data %>% data.frame %>% rownames_to_column("SampID")
div <- read.csv(here("EWG_github_2023","output", "csvs", "inext_div.csv")) %>% dplyr::select(SampID,Estimate) %>% rename(Diversity = Estimate)
rich <- read.csv(here("EWG_github_2023","output", "csvs", "inext_rich.csv")) %>% dplyr::select(SampID,Estimate) %>% rename(Richness = Estimate)

estimates <- left_join(div,rich)
meta_data <- left_join(meta_data, estimates)

mean_diversity <- meta_data %>%
  group_by(Garden, Genotype, Region) %>%
  summarise_at(c("Diversity", "Richness"), mean, na.rm = TRUE)
```
Find mean leaf trait values for each genotype
*This removed two genotypes that didn't have stomata data 
```{r}
leaf_trait_data <- read.csv(here("EWG_github_2023","data", "gardenData", "trait_plus_meta_csvs", "trait_wmeta_allblocks.csv"))

leaf_trait_data %<>%
  dplyr::select(Sample_ID, Garden, Genotype, Region, Population, thick, SLA, LDMC, stomata_length, stomata_count, percent_C_weight,percent_N_weight)

leaf_trait_data %<>% 
  rename(LT = thick) %>%
  rename(SL = stomata_length) %>%
  rename(SD = stomata_count) %>%
  rename(N = percent_N_weight) %>%
  rename(C = percent_C_weight)
         

ltd_avg <- leaf_trait_data %>%
  dplyr::select(Garden, Genotype, LT:N) %>%
  group_by(Garden, Genotype) %>%
  summarise(across(everything(), ~ mean(.x, na.rm = TRUE))) %>%
  na.omit()

ltd_avg <- left_join(ltd_avg, mean_diversity, by = c("Genotype", "Garden")) %>% unique()
```
Relative abundance of F.1 Cladosporium and F.2 - Melampsora
```{r}
phy.clean <- readRDS(here("EWG_github_2023","output", "compiled", "phy.clean.rds"))

phy.clean %<>% phyloseq::filter_taxa(., function(x){sum(x>0) > 0.01*nsamples(.)}, TRUE)
phy.clean <-  transform_sample_counts(phy.clean, function(x) (x / sum(x)))

phy.rust = subset_taxa(phy.clean, Genus=="g__Melampsora")
phy.mycosph = subset_taxa(phy.clean, Genus=="g__Mycosphaerella")

phy.clean_otu <- phy.clean %>% otu_table %>% data.frame

phy.clean_otu %<>%
  rownames_to_column() %>%
  dplyr::select(F.1, F.2, rowname)

meta_data <- phy.clean %>% sample_data %>% data.frame %>% rownames_to_column() %>% dplyr::select(Genotype, Garden, rowname)

otu_abund <- left_join(phy.clean_otu, meta_data, by = "rowname")

otu_abund <- otu_abund %>%
  dplyr::select(Garden, Genotype, F.1, F.2) %>%
  group_by(Garden, Genotype) %>%
  summarise(across(everything(), ~ mean(.x, na.rm = TRUE)))

otu_abund_wdiv <- left_join(otu_abund, ltd_avg, by = c("Genotype", "Garden")) %>% na.omit()

otu_abund_wdiv %<>% mutate_at(c(3:11, 13, 14), scale) 
```

Calculate all Spearman correlations with Bonferroni adjustments
```{r}
library(Hmisc)
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

otu_abund_wdiv_all <- otu_abund_wdiv %>% dplyr::select(-c(Region,Garden,Genotype)) %>% ungroup()
otu_abund_wdiv_all<-otu_abund_wdiv_all[,-1]

otu_res_all<-rcorr(as.matrix(otu_abund_wdiv_all, type="spearman"))

all_corrs <- flattenCorrMatrix(otu_res_all$r, otu_res_all$P)
all_corrs$p.adjusted <- p.adjust(all_corrs$p, method = "fdr")
```
Spearman correlation figure between leaf traits, alpha diversity measures, dominant OTUs by rel abund (F.1 = Cladosporium, F.2 = Melampsora)
```{r}
library(ggstatsplot)
otu_abund_wdiv_all <- otu_abund_wdiv %>% dplyr::select(-c(Region,Genotype)) %>% ungroup()
corr_plot <- grouped_ggcorrmat(
  otu_abund_wdiv_all,
  sig.level = 1, # Recorded sig corr at 0.05 but put sig.level = 1 to remove "X"s from figure
  p.adjust.method = "fdr", 
  type = "n",
  grouping.var = Garden,
  plotgrid.args = list(nrow = 1),
  insig = 'blank',
  pch.col = '#F0E442',
  annotation.args = list(
    tag_levels = "a",
    title = "Relationship between leaf traits and the microbiome across gardens"
  )
  )

corr_plot
ggsave("EWG_github_2023/output/figs/figure6.png")
```